Re-fit

### 2022-06-27 scripts/r/04_refit/refit_tr.Rmd
refit

# libs
library("here")
library("fs")
library("dplyr")
library("tidyr")
library("readr")
library("forcats")
library("ggplot2")
library("stringr")
library("strex")
library("purrr")
library("lme4")
library("brms")
library("broom")
library("broom.mixed")
library("readxl")
library("optimx")

@@

# Functional sequence to get relevant info from Rds files
clean_up <- . %>% 
  tidy(effects = "fixed") %>% 
  suppressWarnings() %>% 
  filter(term %in% c("effect_cattypical", "effect_con")) %>% 
  select(term, estimate, se = std.error)

```
```

@@

# adjective model - pitch
pitch_adj_data = pitchfile %>% filter(word == 9)%>% mutate(type = "adjective")
pitch_noun_data = pitchfile %>% filter(word == 10)%>%mutate(type = "noun")

@@

# standardize
duration_adj_data_temp <- duration_adj_data %>% 
  mutate(duration_s = (duration - mean(duration, na.rm =TRUE)) / sd(duration,  na.rm = TRUE),
  # dummy-code typicality
  typicality = as.factor(typicality),
  target_colour = as.factor(target_colour),
  effect_cat = C(typicality, treatment),
  # sum-code categorical predictors
  target_colour_s = C(target_colour, sum)
  )

duration_noun_data <- duration_noun_data %>% 
  mutate(
         # scale continuous variables
         duration_s = (duration - mean(duration, na.rm = TRUE)) / sd(duration,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour = as.factor(target_colour),
         effect_cat = C(typicality, treatment),
         # sum-code categorical predictors
         target_colour_s = C(target_colour, sum)
  )

pitch_adj_data <- pitch_adj_data %>% 
  mutate(
         # scale continuous variables
         minimum_s = (minimum - mean(minimum, na.rm = TRUE)) / sd(minimum,  na.rm = TRUE),
         maximum_s = (maximum - mean(maximum, na.rm = TRUE)) / sd(maximum,  na.rm = TRUE),
         range_s = (range - mean(range, na.rm = TRUE)) / sd(range,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour = as.factor(target_colour),
         effect_cat = C(typicality, treatment),
         # sum-code categorical predictors
         target_colour_s = C(target_colour, sum)
  )

pitch_noun_data <- pitch_noun_data %>% 
  mutate(
         # scale continuous variables
         minimum_s = (minimum - mean(minimum, na.rm = TRUE)) / sd(minimum,  na.rm = TRUE),
         maximum_s = (maximum - mean(maximum, na.rm = TRUE)) / sd(maximum,  na.rm = TRUE),
         range_s = (range - mean(range, na.rm = TRUE)) / sd(range,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour = as.factor(target_colour),
         effect_cat = C(typicality, treatment),
         # sum-code categorical predictors
         target_colour_s = C(target_colour, sum)
  )

# run brms equivalents
trachyphyllia_lappa_1_duradj_cat <- brm(
  duration_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = duration_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_1_duradj_cat"
  )

trachyphyllia_lappa_2_f0maxadj_cat <- brm(
  maximum_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_2_f0maxadj_cat"
  )

trachyphyllia_lappa_3_f0minadj_cat <- brm(
  minimum_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_3_f0minadj_cat"
  )

trachyphyllia_lappa_4_f0rangeadj_cat <- brm(
  range_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_4_f0rangeadj_cat"
  )

trachyphyllia_lappa_5_durnoun_cat <- brm(
  duration_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = duration_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_5_durnoun_cat"
  )

trachyphyllia_lappa_6_f0maxnoun_cat <- brm(
  maximum_s ~  typicality + target_colour + 
      (typicality + target_colour | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_6_f0maxnoun_cat"
  )

trachyphyllia_lappa_7_f0minnoun_cat <- brm(
  minimum_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_7_f0minnoun_cat"
  )

trachyphyllia_lappa_8_f0rangenoun_cat <- brm(
  range_s ~  typicality + target_colour + 
      (typicality + target_colour  | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_8_f0rangenoun_cat"
  )

@@

# standardize datasets for schwa analyses
MSA_finale <- MSA_finale %>% 
  mutate(
         # scale continuous variables
         f1_time_0_5_s = (f1_time_0_5 - mean(f1_time_0_5, na.rm = TRUE)) / sd(f1_time_0_5,  na.rm = TRUE),
         f2_time_0_5_s = (f2_time_0_5 - mean(f2_time_0_5, na.rm = TRUE)) / sd(f2_time_0_5,  na.rm = TRUE),
         maxIntensity_s = (maxIntensity - mean(maxIntensity, na.rm = TRUE)) / sd(maxIntensity,  na.rm = TRUE),
         meanPitch = as.integer(meanPitch),
         meanPitch_s = (meanPitch - mean(meanPitch, na.rm = TRUE)) / sd(meanPitch,  na.rm = TRUE),
         Target.segment.duration_s = (Target.segment.duration - mean(Target.segment.duration, na.rm = TRUE)) / sd(Target.segment.duration,  na.rm = TRUE),
         Target.word.duration_s = (Target.word.duration - mean(Target.word.duration, na.rm = TRUE)) / sd(Target.word.duration,  na.rm = TRUE),
         Pause.duration_s = (Pause.duration - mean(Pause.duration, na.rm = TRUE)) / sd(Pause.duration,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         controlPitch_s = (controlPitch - mean(controlPitch, na.rm = TRUE)) / sd(controlPitch,  na.rm = TRUE),
         controlIntensity_s = (controlIntensity - mean(controlIntensity, na.rm = TRUE)) / sd(controlIntensity,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = C(as.factor(condition), sum)
  )

# standardize datasets for schwa offglide analyses
MSA_finalschwa <- MSA_finalschwa %>% 
  mutate(
         # scale continuous variables
         f1_time_0_5_s = (f1_time_0_5 - mean(f1_time_0_5, na.rm = TRUE)) / sd(f1_time_0_5,  na.rm = TRUE),
         f2_time_0_5_s = (f2_time_0_5 - mean(f2_time_0_5, na.rm = TRUE)) / sd(f2_time_0_5,  na.rm = TRUE),
         maxIntensity_s = (maxIntensity - mean(maxIntensity, na.rm = TRUE)) / sd(maxIntensity,  na.rm = TRUE),
         meanPitch = as.integer(meanPitch),
         meanPitch_s = (meanPitch - mean(meanPitch, na.rm = TRUE)) / sd(meanPitch,  na.rm = TRUE),
         Target.segment.duration_s = (Target.segment.duration - mean(Target.segment.duration, na.rm = TRUE)) / sd(Target.segment.duration,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         controlPitch_s = (controlPitch - mean(controlPitch, na.rm = TRUE)) / sd(controlPitch,  na.rm = TRUE),
         controlIntensity_s = (controlIntensity - mean(controlIntensity, na.rm = TRUE)) / sd(controlIntensity,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = C(as.factor(condition), sum)
  )

# standardize datasets for schwa + schwa offglide analyses
MSA_both <- MSA_both %>% 
  mutate(
         # scale continuous variables
         DiphDur_s = (DiphDur - mean(DiphDur, na.rm = TRUE)) / sd(DiphDur,  na.rm = TRUE),
         ED_s = (ED - mean(ED, na.rm = TRUE)) / sd(ED,  na.rm = TRUE),
         DurRat_s = (DurRat - mean(DurRat, na.rm = TRUE)) / sd(DurRat,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = C(as.factor(condition), sum)
  )

@@

@@

@@

>>>>

# libs
library("here")
library("fs")
library("dplyr")
library("tidyr")
library("readr")
library("forcats")
library("ggplot2")
library("stringr")
library("strex")
library("purrr")
library("lme4")
library("brms")
library("broom")
library("broom.mixed")
library("readxl")
library("optimx")
library("MuMIn")
library("fda")
library("abind")
library("data.table")
library("magrittr")
library("tidyverse")

# NOTE: one of the many packages interferes with tidyverse, so the select function throws an error!

@@

# Functional sequence to get relevant info from Rds files
clean_up <- . %>% 
  tidy(effects = "fixed") %>% 
  suppressWarnings() %>% 
  filter(term %in% c("effect_cattypical", "effect_con")) %>% 
  dplyr::select(term, estimate, se = std.error)

@@

# adjective model - pitch
pitch_adj_data = pitchfile %>% filter(word == 9) %>% mutate(type = "adjective")
pitch_noun_data = pitchfile %>% filter(word == 10) %>% mutate(type = "noun")

@@

# standardize
duration_adj_data <- duration_adj_data %>% 
  mutate(
    #duration_s = (duration - mean(duration, na.rm = TRUE)) / sd(duration,  na.rm = TRUE),
    # does not work and I don't know why. 
    # dummy-code typicality
    typicality = as.factor(typicality),
    target_colour_s = as.factor(target_colour),
    effect_cat = C(typicality, treatment))

contrasts(duration_adj_data$target_colour_s) <- contr.sum(5)

duration_adj_data$duration_s = (duration_adj_data$duration - mean(duration_adj_data$duration, na.rm = TRUE)) / sd(duration_adj_data$duration,  na.rm = TRUE)

duration_noun_data <- duration_noun_data %>% 
  mutate(
         # scale continuous variables
         # duration_s = (duration - mean(duration, na.rm = TRUE)) / sd(duration,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour_s = as.factor(target_colour),
         effect_cat = C(typicality, treatment)
  )

contrasts(duration_noun_data$target_colour_s) <- contr.sum(5)

duration_noun_data$duration_s = (duration_noun_data$duration - mean(duration_noun_data$duration, na.rm = TRUE)) / sd(duration_noun_data$duration,  na.rm = TRUE)

pitch_adj_data <- pitch_adj_data %>% 
  mutate(
         # scale continuous variables
         # minimum_s = (minimum - mean(minimum, na.rm = TRUE)) / sd(minimum,  na.rm = TRUE),
         # maximum_s = (maximum - mean(maximum, na.rm = TRUE)) / sd(maximum,  na.rm = TRUE),
         # range_s = (range - mean(range, na.rm = TRUE)) / sd(range,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour_s = as.factor(target_colour),
         effect_cat = C(typicality, treatment)
  )

contrasts(pitch_adj_data$target_colour_s) <- contr.sum(5)

pitch_adj_data$minimum_s = (pitch_adj_data$minimum - mean(pitch_adj_data$minimum, na.rm = TRUE)) / sd(pitch_adj_data$minimum,  na.rm = TRUE)
pitch_adj_data$maximum_s = (pitch_adj_data$maximum - mean(pitch_adj_data$maximum, na.rm = TRUE)) / sd(pitch_adj_data$maximum,  na.rm = TRUE)
pitch_adj_data$range_s = (pitch_adj_data$range - mean(pitch_adj_data$range, na.rm = TRUE)) / sd(pitch_adj_data$range,  na.rm = TRUE)

pitch_noun_data <- pitch_noun_data %>% 
  mutate(
         # scale continuous variables
         #minimum_s = (minimum - mean(minimum, na.rm = TRUE)) / sd(minimum,  na.rm = TRUE),
         #maximum_s = (maximum - mean(maximum, na.rm = TRUE)) / sd(maximum,  na.rm = TRUE),
         #range_s = (range - mean(range, na.rm = TRUE)) / sd(range,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour_s = as.factor(target_colour),
         effect_cat = C(typicality, treatment),
         # sum-code categorical predictors
         # target_colour_s = C(target_colour, sum, 5)
  )

# sum-code categorical predictors
contrasts(pitch_noun_data$target_colour_s) <- contr.sum(5)

pitch_noun_data$minimum_s = (pitch_noun_data$minimum - mean(pitch_noun_data$minimum, na.rm = TRUE)) / sd(pitch_noun_data$minimum,  na.rm = TRUE)
pitch_noun_data$maximum_s = (pitch_noun_data$maximum - mean(pitch_noun_data$maximum, na.rm = TRUE)) / sd(pitch_noun_data$maximum,  na.rm = TRUE)
pitch_noun_data$range_s = (pitch_noun_data$range - mean(pitch_noun_data$range, na.rm = TRUE)) / sd(pitch_noun_data$range,  na.rm = TRUE)

# run brms equivalents
trachyphyllia_lappa_1_duradj_cat <- brm(
  duration_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = duration_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_1_duradj_cat"
  )

trachyphyllia_lappa_2_f0maxadj_cat <- brm(
  maximum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_2_f0maxadj_cat"
  )

trachyphyllia_lappa_3_f0minadj_cat <- brm(
  minimum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) +  
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_3_f0minadj_cat"
  )

trachyphyllia_lappa_4_f0rangeadj_cat <- brm(
  range_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_4_f0rangeadj_cat"
  )

# refitted with more samples because of the interaction
trachyphyllia_lappa_5_durnoun_cat <- brm(
  duration_s ~  effect_cat * target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = duration_noun_data, 
  iter = 4000,
  control = list(adapt_delta = 0.9999),
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_5_durnoun_cat"
  )

trachyphyllia_lappa_6_f0maxnoun_cat <- brm(
  maximum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_6_f0maxnoun_cat"
  )

trachyphyllia_lappa_7_f0minnoun_cat <- brm(
  minimum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) +  
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_7_f0minnoun_cat"
  )

trachyphyllia_lappa_8_f0rangenoun_cat <- brm(
  range_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_8_f0rangenoun_cat"
  )

@@

# standardize datasets for schwa analyses
MSA_finale <- MSA_finale %>% 
  mutate(
         # scale continuous variables
         f1_time_0_5_s = (f1_time_0_5 - mean(f1_time_0_5, na.rm = TRUE)) / sd(f1_time_0_5,  na.rm = TRUE),
         f2_time_0_5_s = (f2_time_0_5 - mean(f2_time_0_5, na.rm = TRUE)) / sd(f2_time_0_5,  na.rm = TRUE),
         maxIntensity_s = (maxIntensity - mean(maxIntensity, na.rm = TRUE)) / sd(maxIntensity,  na.rm = TRUE),
         meanPitch = as.integer(meanPitch),
         meanPitch_s = (meanPitch - mean(meanPitch, na.rm = TRUE)) / sd(meanPitch,  na.rm = TRUE),
         Target.segment.duration_s = (Target.segment.duration - mean(Target.segment.duration, na.rm = TRUE)) / sd(Target.segment.duration,  na.rm = TRUE),
         Target.word.duration_s = (Target.word.duration - mean(Target.word.duration, na.rm = TRUE)) / sd(Target.word.duration,  na.rm = TRUE),
         Pause.duration_s = (Pause.duration - mean(Pause.duration, na.rm = TRUE)) / sd(Pause.duration,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         controlPitch_s = (controlPitch - mean(controlPitch, na.rm = TRUE)) / sd(controlPitch,  na.rm = TRUE),
         controlIntensity_s = (controlIntensity - mean(controlIntensity, na.rm = TRUE)) / sd(controlIntensity,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = as.factor(condition),
         condition_s = C(condition_s, sum)
  )

# standardize datasets for schwa offglide analyses
MSA_finalschwa <- MSA_finalschwa %>% 
  mutate(
         # scale continuous variables
         f1_time_0_5_s = (f1_time_0_5 - mean(f1_time_0_5, na.rm = TRUE)) / sd(f1_time_0_5,  na.rm = TRUE),
         f2_time_0_5_s = (f2_time_0_5 - mean(f2_time_0_5, na.rm = TRUE)) / sd(f2_time_0_5,  na.rm = TRUE),
         maxIntensity_s = (maxIntensity - mean(maxIntensity, na.rm = TRUE)) / sd(maxIntensity,  na.rm = TRUE),
         meanPitch = as.integer(meanPitch),
         meanPitch_s = (meanPitch - mean(meanPitch, na.rm = TRUE)) / sd(meanPitch,  na.rm = TRUE),
         Target.segment.duration_s = (Target.segment.duration - mean(Target.segment.duration, na.rm = TRUE)) / sd(Target.segment.duration,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         controlPitch_s = (controlPitch - mean(controlPitch, na.rm = TRUE)) / sd(controlPitch,  na.rm = TRUE),
         controlIntensity_s = (controlIntensity - mean(controlIntensity, na.rm = TRUE)) / sd(controlIntensity,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = as.factor(condition),
         condition_s = C(condition_s, sum)
  )

# standardize datasets for schwa + schwa offglide analyses
MSA_both <- MSA_both %>% 
  mutate(
         # scale continuous variables
         DiphDur_s = (DiphDur - mean(DiphDur, na.rm = TRUE)) / sd(DiphDur,  na.rm = TRUE),
         ED_s = (ED - mean(ED, na.rm = TRUE)) / sd(ED,  na.rm = TRUE),
         DurRat_s = (DurRat - mean(DurRat, na.rm = TRUE)) / sd(DurRat,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = as.factor(condition),
         condition_s = C(condition_s, sum)
  )

@@

```{r pervagor_meeki_setup}

data = read.csv("./data/analyses/tr/pervagor_meeki/Data/ManySpeech_Praat_merged_data.csv", header=T)

data$Duration_ms = data$Duration*1000
data_to_analyze <- data

########## SET CONTRASTS ############

###### set sum contrasts (compare to grand mean)
data_to_analyze$typicality = factor(data_to_analyze$typicality)

contrasts(data_to_analyze$typicality) = contr.treatment(3)

########### REMOVE OUTLIERS ###########

# TRIMMING >3 SD AND <3 SD, by SPEAKER

########### MEAN PITCH (semitones) ##################
data_meanf0_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Mean_f0_st), stdev = sd(Mean_f0_st)) %>%  
  filter(Mean_f0_st <= (3*stdev)+avg & 
           Mean_f0_st >= avg-(3*stdev)) %>%  
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

########### MAX PITCH (semitones) ##################
data_maxf0_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Max_f0_st), stdev = sd(Max_f0_st)) %>%  
  filter(Max_f0_st <= (3*stdev)+avg & 
           Max_f0_st >= avg-(3*stdev)) %>%  
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

########### INTENSITY (Praat measurements) ##################
data_int_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Overall_Intensity), stdev = sd(Overall_Intensity)) %>%  
  filter(Overall_Intensity <= (3*stdev)+avg & 
           Overall_Intensity >= avg-(3*stdev)) %>% 
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

########### DURATION ##################
data_dur_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Duration_ms), stdev = sd(Duration_ms)) %>%  
  filter(Duration_ms <= (3*stdev)+avg & #slow RTs greater than 3 SD
           Duration_ms >= avg-(3*stdev)) %>%  #fast RTs less than 3 SD
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

##################### MODELS ############################

#set which contrast is dropped from the models

#run models twice:
#base = 3: drop typical
#base = 2: drop medium
base_n = 3

########### MEAN PITCH (semitones) ##################

contrasts(data_meanf0_trim$typicality) = contr.treatment(3,base=base_n)
contrasts(data_meanf0_trim$typicality)

@@

```{r pervagor_meeki_models}

# mean F0
  #NOUN MODEL
  summary( meanf0_mod <- lmer (Mean_f0_st ~ typicality
                               + (typicality | speaker),
                               data = data_meanf0_trim[data_meanf0_trim$word_type == "target_name",],
                               control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( meanf0_mod <- lmer (Mean_f0_st ~ typicality
                               + (1 | speaker)
                               + (1 | target_word),
                               data = data_meanf0_trim[data_meanf0_trim$word_type == "target_colour",],
                               control=lmerControl(optimizer="bobyqa")))

# max F0
   #NOUN MODEL
   summary( maxf0_mod <- lmer (Max_f0_st ~ typicality
                              + (1 | speaker),
                              data = data_maxf0_trim[data_maxf0_trim$word_type == "target_name",],
                              control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( maxf0_mod <- lmer (Max_f0_st ~ typicality
                              + (1 | speaker)
                              + (1 | target_word),
                              data = data_maxf0_trim[data_maxf0_trim$word_type == "target_colour",],
                              control=lmerControl(optimizer="bobyqa")))

# int
  #NOUN MODEL
  summary( int_mod <- lmer (Overall_Intensity ~ typicality
                            + (1 | speaker),
                            data = data_int_trim[data_int_trim$word_type == "target_name",],
                            control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( int_mod <- lmer (Overall_Intensity ~ typicality
                            + (1 | speaker)
                            + (typicality | target_word),
                            data = data_int_trim[data_int_trim$word_type == "target_colour",],
                            control=lmerControl(optimizer="bobyqa")))

# duration  
  #NOUN MODEL
  summary( dur_mod <- lmer (Duration_ms ~ typicality
                            + (1 | speaker),
                            data = data_dur_trim[data_dur_trim$word_type == "target_name",],
                            control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( dur_mod <- lmer (Duration_ms ~ typicality
                            + (1 | speaker)
                            + (typicality | target_word),
                            data = data_dur_trim[data_dur_trim$word_type == "target_colour",],
                            control=lmerControl(optimizer="bobyqa")))
```

```{r pervagor_meeki_refitting}

# scale
data_int_trim <- data_int_trim %>% 
  mutate(
    Overall_Intensity_s = (Overall_Intensity - mean(Overall_Intensity, na.rm = TRUE)) / sd(Overall_Intensity,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

data_meanf0_trim <- data_meanf0_trim %>% 
  mutate(
    Mean_f0_st_s = (Mean_f0_st - mean(Mean_f0_st, na.rm = TRUE)) / sd(Mean_f0_st,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

data_maxf0_trim <- data_maxf0_trim %>% 
  mutate(
    Max_f0_st_s = (Max_f0_st - mean(Max_f0_st, na.rm = TRUE)) / sd(Max_f0_st,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

data_dur_trim <- data_dur_trim %>% 
  mutate(
    Duration_ms_s = (Duration_ms - mean(Duration_ms, na.rm = TRUE)) / sd(Duration_ms,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

pervagor_meeki_1_intnoun_cat <- brm(
  Overall_Intensity_s ~ 
         effect_cat + 
         (1|speaker),
  seed = 111,
  data = data_int_trim[data_int_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_1_intnoun_cat"
  )

pervagor_meeki_2_f0meannoun_cat <- brm(
  Mean_f0_st_s ~ 
         effect_cat + 
         + (effect_cat | speaker),
  seed = 111,
  data = data_meanf0_trim[data_meanf0_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_2_f0meannoun_cat"
  )

pervagor_meeki_3_f0maxnoun_cat <- brm(
  Max_f0_st_s ~ 
         effect_cat + 
         + (1| speaker),
  seed = 111,
  data = data_maxf0_trim[data_maxf0_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_3_f0maxnoun_cat"
  )

pervagor_meeki_4_durnoun_cat <- brm(
  Duration_ms_s ~ 
         effect_cat + 
         + (1| speaker),
  seed = 111,
  data = data_dur_trim[data_dur_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_4_durnoun_cat"
  )

pervagor_meeki_5_intadj_cat <- brm(
  Overall_Intensity_s ~ 
         effect_cat 
         + (1 | speaker)
         + (effect_cat | target_word),
  seed = 111,
  data = data_int_trim[data_int_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_5_intadj_cat"
  )

pervagor_meeki_6_f0meanadj_cat <- brm(
  Mean_f0_st_s ~ 
         effect_cat + 
         + (1 | speaker)
         + (1 | target_word),
  seed = 111,
  data = data_meanf0_trim[data_meanf0_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_6_f0meanadj_cat"
  )

pervagor_meeki_7_f0maxadj_cat <- brm(
  Max_f0_st_s ~ 
         effect_cat + 
         + (1 | speaker)
         + (1 | target_word),
  seed = 111,
  data = data_maxf0_trim[data_maxf0_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_7_f0maxadj_cat"
  )

pervagor_meeki_8_duradj_cat <- brm(
  Duration_ms_s ~ 
         effect_cat + 
          + (1 | speaker)
          + (effect_cat | target_word),
  seed = 111,
  data = data_dur_trim[data_dur_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_8_duradj_cat"
  )

@@

```{r procambarus_maculosus_setup}

getPCscores <- function(fpcaObj) {
  if (is.matrix(fpcaObj$scores)) {
    fpcaObj$scores
  } else if (is.array(fpcaObj$scores) & length(dim(fpcaObj$scores)) == 3) {
    apply(fpcaObj$scores, 1:2, sum)
  } else {
    stop("getPCscores: invalid input fpcaObj")
  }
}

##### load data #####

data <- read.delim("./data/analyses/tr/selectedData.txt", header = T, sep = "") %>% 
  as_tibble() # %>% 
  # filter(sex == "female") # only run on female data!

data = read.csv("./data/analyses/tr/pervagor_meeki/Data/ManySpeech_Praat_merged_data.csv", header=T)

info <- data %>% 
  select(sl_rowIdx, bundle, vowel, speaker, sex, typicality, word) %>% 
  unique()

##### FPCA #####

# set parameters
nKnots <- 6
lambda <- 1e-8
Lfdobj <- 2
nOrder <- 2 + Lfdobj
nBasis <- nKnots + nOrder - 2

# create a basis for B-splines
basis <- create.bspline.basis(rangeval = c(0, 1), nbasis = nBasis, norder = nOrder)

# define functional parameter object (will be used to estimate functional parameters later)
fdParObj <- fdPar(fdobj = basis, Lfdobj = Lfdobj, lambda = lambda)

coef <- data %>%
  group_by(sl_rowIdx) %>%
  summarise(coef = c(smooth.basis(argvals = times_norm, y = F0, fdParObj)$fd$coefs),
            coefId = seq_len(nBasis)) %>%
  ungroup() %>%
  pivot_wider(names_from = coefId, values_from = coef) %>%
  select(-sl_rowIdx) %>% 
  abind(along = 2) %>%
  aperm(c(2,1))

# check class and dimension
coef %>% class() # array
coef %>% dim()   # 8 842; 8 671

# create functional data object
y_fd <- fd(coef = coef, basisobj = basis)

# this is the actual FPCA
nPCs <- 3
y_pcafd <- pca.fd(y_fd, nharm = nPCs, fdParObj)
# saveRDS(y_pcafd, file = "./pcafd.rds")

y_pcafd$varprop 
# both sexes: 0.92432242 0.04662226 0.01509360
# only female: 0.87031521 0.08118398 0.02410822

info %<>% 
  cbind(y_pcafd %>% getPCscores) %>%
  rename(s1 = `1`, s2 = `2`, s3 = `3`) %>% 
  as_tibble()
# write.table(info, "./PCscores.txt")

##### interpretation #####

tx <- seq(1/1000, 1, by = 1/1000)

curves <- expand.grid(time = tx,
                      PC = 1:3,
                      perturbation = c(-1, 0, 1)
) %>% setDT

sd_score <- info %>% summarise_at(vars(s1:s3), sd) %>% as.numeric()

curves[, value := (y_pcafd$meanfd$coefs[,1] + perturbation * sd_score[PC] * y_pcafd$harmonics$coefs[,PC]) %>% 
         fd(., basis) %>%
         eval.fd(tx, .),
       by = .(PC, perturbation)]

reconstruct <- curves %>% mutate(per = ifelse(perturbation == 0, "mu[F0](t)",
                                              ifelse(perturbation == 1, "mu[F0](t) + sigma[s[k]] %.% PCk[F0](t)",
                                                     "mu[F0](t) - sigma[s[k]] %.% PCk[F0](t)")),
                                 k = case_when(PC == 1 ~ "k == 1",
                                               PC == 2 ~ "k == 2",
                                               PC == 3 ~ "k == 3"))

reconstruct$per <- factor(reconstruct$per)

levels(reconstruct$per) = list("mu[F0](t) - sigma[s[k]] %.% PCk[F0](t)" = "0",
                               "mu[F0](t)" = "1",
                               "mu[F0](t) + sigma[s[k]] %.% PCk[F0](t)" = "2")
